---
title: "192_MP4"
author: "Dasom An & Janell Lin"
date: "May 11th, 2018"
output: 
 html_document:
   code_folding: hide
---

# Gender Bias Against Directors in Film Field?

According to a paper "Gender Stereotypes Are Alive, Well, and Busy Producing Workplace Discrimination" published by [Cambridge University](https://www.cambridge.org/core/journals/industrial-and-organizational-psychology/article/gender-stereotypes-are-alive-well-and-busy-producing-workplace-discrimination/79B1C722310ABBBE8D0EA90157689B97), females have been a victim of gender bias in workplace. Recently, there have been many movements related to gender issues like "MeToo" against sexual harassment and assult and increased women's vocie against wage discrimination or glass ceiling in their workplace.

Especially in film field, gender discrimination and stereotype have been hot issues. Our hypothesis is that there has been implicit prejudice that a female director would make a low quality movie compared to one by a male director. We sought to determine if female and male directors are well diversified in film field and if there is an association between gender of directors and the quality of movie. 

We used **imdb** database source from Smith College and extracted recent 10-year best movies based on rating and their revenue. We set the production year from 2008 to 2018 in order to 



```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(mdsr)
library(RMySQL)
db <- dbConnect_scidb(dbname = "imdb")
db %>%
  dbGetQuery("SELECT * FROM kind_type;")
```

```{r, message=FALSE, warning=FALSE}
# Extracting 100 best movies info from database

sql <- "
SELECT t.id, t.title, t.production_year, mii.info AS votes, mii2.info AS rating
FROM title t
JOIN movie_info_idx mii ON mii.movie_id = t.id
JOIN movie_info_idx mii2 ON mii2.movie_id = t.id
WHERE t.production_year >= 2008
AND t.kind_id = 1
AND mii.info_type_id = 100
AND mii2.info_type_id = 101
AND mii.info > 100000
ORDER BY mii2.info desc;
"
best_movies <- db %>%
  dbGetQuery(sql)
head(best_movies)

best_movies[!duplicated(best_movies[,c("title")]), ]

best_movies100 <- best_movies %>%
  head(100)
```

```{r, message=FALSE, warning=FALSE}
# Extracting directors info from database

sql_1 <- "SELECT n.name, ci.role_id, person_id,  movie_id, n.gender 
FROM cast_info ci
JOIN name n ON n.id = ci.person_id
WHERE role_id = '8';
"
people_info <- db %>%
  dbGetQuery(sql_1)
head(people_info)

all_info<- people_info %>%
  select(name, person_id, gender, movie_id)%>%
  group_by(name, person_id, gender, movie_id)%>%
  summarise(num=n()) 
```

```{r, message=FALSE, warning=FALSE}
# Join movie_id with best_movies100 for their movie_id

sql_2 <- "
SELECT movie_id, title, production_year, id 
FROM aka_title;
"
movie_id <- db %>%
  dbGetQuery(sql_2)

movie_id1<- movie_id %>%
  filter(production_year>=2008)
 group_by (movie_id, title) %>%
  summarise (num = n())

head(movie_id)

# Joining best 100 movies with their movie_id
movie_id2 <- movie_id1 %>%
  inner_join(best_movies, by = c("title"= "title", "production_year" = "production_year")) 

movie_id100<-movie_id2 %>%
  head (100)

movie_id100[!duplicated(movie_id100[,c("title")]), ]
```


```{r, message=FALSE, warning=FALSE}
# Extracting 100 best movies gross info from database

sql_gross <- "
SELECT t.id, t.title, mi.info AS gross_info
FROM title t
JOIN movie_info mi ON mi.movie_id = t.id;
"

gross <- db %>%
dbGetQuery(sql_gross)
head(gross)


dbGetQuery(db, "SELECT * FROM info_type WHERE info = 'gross';")

dbGetQuery(db, "
SELECT * FROM info_type 
WHERE info LIKE '%gross%';
")

movie_gross<- dbGetQuery(db, "
SELECT t.id, t.title, mi.info AS gross_info, movie_id, production_year
FROM title t
JOIN movie_info mi ON mi.movie_id = t.id
WHERE mi.info_type_id = 107
")

movie_gross_1 <- movie_gross %>%
  group_by (movie_id)%>%
  filter(grepl("USA", gross_info))%>%
  summarise(max(gross_info))

movie_gross100<- movie_gross_1%>%
  inner_join(movie_id2, by= c("movie_id"="movie_id"))%>%
  head(100)

movie_test <- movie_gross100 %>%
  mutate(Revenue=as.numeric(gross_info))
  
```

```{r, message=FALSE, warning=FALSE}
# Extracting gross info for best movies

sql_gross <- "
SELECT t.id, t.title, mi.info AS gross_info
FROM title t
JOIN movie_info mi ON mi.movie_id = t.id;
"

gross <- db %>%
dbGetQuery(sql_gross)
head(gross)


dbGetQuery(db, "SELECT * FROM info_type WHERE info = 'gross';")

dbGetQuery(db, "
SELECT * FROM info_type 
WHERE info LIKE '%gross%';
")

movie_gross<- dbGetQuery(db, "
SELECT t.id, t.title, mi.info AS gross_info, movie_id, production_year
FROM title t
JOIN movie_info mi ON mi.movie_id = t.id
WHERE mi.info_type_id = 107
")

movie_gross_1 <- movie_gross %>%
  group_by (movie_id)%>%
  filter(grepl("USA", gross_info))%>%
  summarise(max(gross_info))

movie_gross100<- movie_gross_1%>%
  inner_join(movie_id2, by= c("movie_id"="movie_id"))%>%
  head(100)

names(movie_gross100)[names(movie_gross100)=="max(gross_info)"]<-"gross_info" 
```

```{r, message=FALSE, warning=FALSE}
# Joining directors info with best movies

movie_people_info <- movie_gross100 %>%
  inner_join(all_info, by= c("movie_id"="movie_id"))

MP_info <- movie_people_info %>%
  select(title,movie_id, gross_info, production_year,gender,rating,votes, name)

```

```{r}
# Visualizing findings
library(scales)

ggplot(data=MP_info, aes(x=gross_info, y=rating)) +
  geom_point(aes(color=gender)) +
  labs(title = "Quality of Best 100 Movies by Male and Female Directors", x = "Revenue($)", y = "Rating")
 
--------
  CHECKING THROUGH HERE !! 
   
# We need to set the revenue units first to set x-axis to annotate 
  scale_y_discrete(labels = c("4.5", "5.0", "5.5", "6.0", "6.5", "7.0", "7.5", "8.0", "8.5"), limits = c(4.0, 9.0), breaks = seq(4.5,8.5,0.5)) +
    scale_x_discrete(labels = c("25M", "50M", "75M", "100M"), limits = c(0,105000000), breaks = seq(25000000,100000000, 25000000)) 
  

# Examples of annotating and so on from babynames hw
annotate("segment", x=1940, xend=1948, y=0.5, yend=0.5, color="black")+
annotate("segment", x=1940, xend=1940, y=0.4, yend=0.5, color="black") +

labels =c("1940", "'60", "'80", "2000" ), limits =c(1930,2012) ,breaks =seq(1940,2000,20


                                                                            
# Testing changing gross things to number                                                                              
movie_test <- movie_gross100 %>%
  mutate(Revenue=as.numeric(gross_info))                                                                              
                                                                              
# Example from mp3                                                                               
mutate (Difficulty= if_else(as.numeric(numeric_rating)<10, "Easy", ifelse(as.numeric(numeric_rating)<30, "Moderate", "Hard")))


```


According to our data, 5% of best 100 movies is directed by female director, which means gender diversity is not satisfied. 
In terms of the movie quality, we could not find significant evidence to prove that the effect of female directors on film performance would be less than the effect of male directors.  
In our graph, red points (movies by female directors) are situated on high revenue and high movie rating. 



Through our database querying, we conclude that gender diversity would help to improve the film field. Even if there are a few female directors who made big hit, their impact is not small.  

Not only through our data findings, there have been many recent papers to show that gender diversity would be good for better environment and more prosperity in film field. According to a paper "Female Directors And Firm Performance" published by [Kwik Kian Gie School of Business](https://www.researchgate.net/publication/317183034_FEMALE_DIRECTORS_AND_FIRM_PERFORMANCE_EVIDENCE_FROM_THE_UK_LISTED_FIRMS?enrichId=rgreq-28cf48989bff249ec4e0ef497945aa8b-XXX&enrichSource=Y292ZXJQYWdlOzMxNzE4MzAzNDtBUzo0OTkxNjcwMzU0MzI5NjBAMTQ5NjAyMjEwMzQ3NQ%3D%3D&el=1_x_2&_esc=publicationCoverPdf), gender diversity helps to improve creativity, innovation, and decision making for better problem solvings, therefore, causes positive results on film market.



**LINKS**

GitHub repository[^1]

Gender Stereotypes Are Alive, Well, and Busy Producing Workplace Discrimination, Cambridge University [^2]

Female Directors And Firm Performance, Kwik Kian Gie School of Business [^3]

[^1]: 
[^2]: https://www.cambridge.org/core/journals/industrial-and-organizational-psychology/article/gender-stereotypes-are-alive-well-and-busy-producing-workplace-discrimination/79B1C722310ABBBE8D0EA90157689B97
[^3]: https://www.researchgate.net/publication/317183034_FEMALE_DIRECTORS_AND_FIRM_PERFORMANCE_EVIDENCE_FROM_THE_UK_LISTED_FIRMS?enrichId=rgreq-28cf48989bff249ec4e0ef497945aa8b-XXX&enrichSource=Y292ZXJQYWdlOzMxNzE4MzAzNDtBUzo0OTkxNjcwMzU0MzI5NjBAMTQ5NjAyMjEwMzQ3NQ%3D%3D&el=1_x_2&_esc=publicationCoverPdf 
Female Directors And Firm Performance, Kwik Kian Gie School of Business 


